---
output:
  word_document: default
  html_notebook:
    df_print: paged
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# Setup knitr options
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 5
)
```

## Load Libraries and Data

```{r load-libraries}
# Install and load required packages
pkgs <- c("readr", "dplyr", "tidyr", "ggplot2", "gridExtra", "ARTool", "boot", "reshape2", "viridis", "knitr", "kableExtra")
for (p in pkgs) {
  if (!requireNamespace(p, quietly = TRUE)) install.packages(p)
  library(p, character.only = TRUE)
}

# Read merged results CSV and prepare factors
df <- readr::read_csv("merged_results.csv") %>%
  dplyr::mutate(
    size  = factor(paste0(size, "B"), levels = c("1B","3B","11B")),
    quant = factor(quant, levels = c("bf16","4bit"))
  )

```

## 1. Assumption Checks & Statistical Approach

### 1.1 Shapiro–Wilk Normality Test

```{r shapiro-test}
# Subset 4-bit fine-tuned adapters only
df_ft <- dplyr::filter(df, quant == "4bit", !is.na(seed))

# Fit one-way ANOVA and extract residuals
aov_model <- stats::aov(accuracy ~ size, data = df_ft)
resid_ft <- stats::residuals(aov_model)

# Shapiro–Wilk test for normality of residuals
stats::shapiro.test(resid_ft)
```

### 1.2 Kruskal–Wallis One-Way Test

```{r kruskal-test}
# Non-parametric one-way test on accuracy by size
stats::kruskal.test(accuracy ~ size, data = df_ft)
```
### 1.3 Pairwise Wilconxon Test
```{r posthoc-wilcox}
# Pairwise comparisons of accuracy between adapter sizes
stats::pairwise.wilcox.test(
  x = df_ft$accuracy,
  g = df_ft$size,
  p.adjust.method = "bonferroni"
)
```

### 1.4 Aligned-Rank Transform Two-Way ANOVA (ART)

```{r art-anova}
# Fit ART model for main effects and interaction
art_model <- ARTool::art(accuracy ~ size * quant, data = df)
stats::anova(art_model)
```

## 2. Chart Generation

### 2.1 Bootstrap Confidence Interval Function

```{r bootstrap-function}
# Function to compute 99% bootstrap CI manually
bootstrap_ci <- function(x, n = 2000, alpha = 0.01) {
  # Draw bootstrap samples of means
  boot_means <- replicate(n, mean(sample(x, replace = TRUE)))
  # Compute percentile CI
  ci <- quantile(boot_means, probs = c(alpha/2, 1 - alpha/2))
  return(ci)
}
```

### 2.2 Figure 1: Mean Accuracy & F1 with 99% Bootstrap CIs for 4BIT Fine-tuned Models

```{r fig1-accuracy-f1}
# Compute manual bootstrap-percentile CIs to avoid constant-data errors
boot_ci <- function(x, R=2000, alpha=0.01) {
  boots <- replicate(R, mean(sample(x, replace=TRUE)))
  qs   <- quantile(boots, probs=c(alpha/2, 1-alpha/2))
  return(c(lower=qs[1], upper=qs[2]))
}

# Summarize means + CIs for 4-bit adapters only
data_summary <- df %>%
  filter(quant == "4bit", !is.na(seed)) %>%
  group_by(size) %>%
  summarise(
    acc_mean    = mean(accuracy),
    acc_ci_low  = boot_ci(accuracy)["lower"],
    acc_ci_high = boot_ci(accuracy)["upper"],
    f1_mean     = mean(f1),
    f1_ci_low   = boot_ci(f1)["lower"],
    f1_ci_high  = boot_ci(f1)["upper"],
    .groups      = "drop"
  )

# Plot Accuracy with error bars
p1 <- ggplot(data_summary, aes(x = size, y = acc_mean)) +
  geom_col(fill = "#66c2a5") +
  geom_errorbar(aes(ymin = acc_ci_low, ymax = acc_ci_high),
                width = 0.2) +
  labs(y = "Mean Accuracy ±99% CI", x = "Model Size") +
  theme_minimal()

# Plot F1 with error bars
p2 <- ggplot(data_summary, aes(x = size, y = f1_mean)) +
  geom_col(fill = "#fc8d62") +
  geom_errorbar(aes(ymin = f1_ci_low, ymax = f1_ci_high),
                width = 0.2) +
  labs(y = "Mean F1 ±99% CI", x = "Model Size") +
  theme_minimal()

# Arrange side by side
gridExtra::grid.arrange(p1, p2, ncol = 2)

data_summary

```

### 2.3 Figure 2: Seed-Rep Variability (4-bit FT) Figure 2: Seed-Rep Variability (4-bit FT)

```{r fig2-variability}
# Boxplots for accuracy and F1 by size
gg1 <- ggplot(data=df_ft, aes(x = size, y = accuracy)) +
  geom_boxplot(fill = '#66c2a5') +
  labs(y = 'Accuracy', x = 'Model Size') +
  ggtitle('Accuracy Variability (4-bit FT)') +
  theme_minimal()

gg2 <- ggplot(data=df_ft, aes(x = size, y = f1)) +
  geom_boxplot(fill = '#fc8d62') +
  labs(y = 'F1', x = 'Model Size') +
  ggtitle('F1 Variability (4-bit FT)') +
  theme_minimal()

gridExtra::grid.arrange(gg1, gg2, ncol = 1)

summary_tbl <- df_ft %>%
  group_by(size) %>%
  summarize(
    acc_min    = min(accuracy),
    acc_q1     = quantile(accuracy, 0.25),
    acc_median = median(accuracy),
    acc_q3     = quantile(accuracy, 0.75),
    acc_max    = max(accuracy),
    f1_min     = min(f1),
    f1_q1      = quantile(f1, 0.25),
    f1_median  = median(f1),
    f1_q3      = quantile(f1, 0.75),
    f1_max     = max(f1),
    .groups = "drop"
  )

# Render as a styled table
summary_tbl %>%
  kable(
    format  = "html",
    digits  = 3,
    caption = "Table: Accuracy & F1 Summary by Model Size (4-bit Fine-Tuning)"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width        = FALSE
  ) %>%
  column_spec(1, bold = TRUE) %>%              # Bold the size column
  column_spec(2:6, width = "1.2cm") %>%        # Set width for accuracy stats
  column_spec(7:11, width = "1.2cm")           # Set width for F1 stats
  
```

### 2.4 Figure 3: Descriptive Interaction Plot

```{r fig3-interaction}
# Mean accuracy by size & quant
# Prepare a grouping variable
df_plot <- df %>%
  mutate(
    group = case_when(
      is.na(seed) & quant == "bf16" ~ "Base BF16",
      is.na(seed) & quant == "4bit" ~ "Base 4-bit",
      !is.na(seed) & quant == "4bit" ~ "FT 4-bit"
    )
  )

# Plot all three
ggplot(df_plot, aes(x = size, y = accuracy, color = group, group = group)) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  stat_summary(fun = mean, geom = "point", size = 2) +
  labs(
    title = "Mean Accuracy by Size and Model Type",
    x = "Model Size",
    y = "Mean Accuracy",
    color = NULL
  ) +
  theme_minimal()

# Compute mean accuracy summary
summary_tbl2 <- df_plot %>%
  group_by(size, group) %>%
  summarize(
    mean_accuracy = mean(accuracy, na.rm = TRUE),
    .groups       = "drop"
  )

summary_tbl2 %>%
  kable(
    format  = "html",
    digits  = 3,
    caption = "Mean Accuracy by Model Size and Type"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE
  ) %>%
  column_spec(1, bold = TRUE) %>%                # Bold the size column
  column_spec(2, width = "3cm") %>%              # Make group column wider
  column_spec(3, width = "2.5cm")                # Set width for mean_accuracy

```

### 2.5 Figure 4: Median Accuracy with IQR

```{r fig4-median-iqr}
# Median accuracy and IQR by size & model group
df_plot <- df %>%
  dplyr::mutate(
    group = dplyr::case_when(
      is.na(seed) & quant == "bf16" ~ "Base BF16",
      is.na(seed) & quant == "4bit" ~ "Base 4-bit",
      !is.na(seed) & quant == "4bit" ~ "FT 4-bit"
    )
  )

df_plot %>%
  group_by(size, group) %>%
  summarise(
    med   = median(accuracy),
    q1    = quantile(accuracy, 0.25),
    q3    = quantile(accuracy, 0.75),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = size, y = med, ymin = q1, ymax = q3, color = group)) +
    geom_point(size = 3, position = position_dodge(width = 0.4)) +
    geom_errorbar(
      position = position_dodge(width = 0.4),
      width = 0.2
    ) +
    labs(
      title = "Median Accuracy and IQR by Model Group",
      x     = "Model Size",
      y     = "Median Accuracy",
      color = NULL
    ) +
    theme_minimal()


summary_tbl3 <- df_plot %>%
  group_by(size, group) %>%
  summarise(
    med = median(accuracy, na.rm = TRUE),
    q1  = quantile(accuracy, 0.25, na.rm = TRUE),
    q3  = quantile(accuracy, 0.75, na.rm = TRUE),
    .groups = "drop"
  )


summary_tbl3 %>%
  kable(
    format  = "html",
    digits  = 3,
    caption = "Median Accuracy and IQR by Model Size and Group"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE
  ) %>%
  column_spec(1, bold = TRUE) %>%               # Bold the size column
  column_spec(2, width = "3cm") %>%             # Wider for group names
  column_spec(3:5, width = "2cm")               # Set width for med, q1, q3
```

### 2.6 Figure 5: Residual Diagnostics

```{r fig5-residuals}
# QQ-plot and Residuals vs Fitted for 4-bit adapters
par(mfrow = c(1, 2))
# QQ-plot
qqnorm(resid_ft)
qqline(resid_ft)
# Residuals vs Fitted
plot(fitted(aov_model), resid_ft,
     xlab = "Fitted Values", ylab = "Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, lty = 2)
# Reset graphics
par(mfrow = c(1, 1))
```

### 2.7 Figure 6: Resource Usage by Size & Quant

```{r fig6-resources}
# GPU & CPU summary by model group
res_summary <- df_plot %>%
  group_by(size, group) %>%
  summarise(
    gpu = mean(gpu_peak_mem_mb),
    cpu = mean(cpu_rss_mb),
    .groups = "drop"
  )

p3 <- ggplot(res_summary, aes(x = size, y = gpu, fill = group)) +
  geom_col(position = position_dodge(width = 0.8)) +
  labs(
    y = "GPU Peak Memory (MB)",
    x = "Model Size",
    fill = NULL
  ) +
  theme_minimal()

p4 <- ggplot(res_summary, aes(x = size, y = cpu, fill = group)) +
  geom_col(position = position_dodge(width = 0.8)) +
  labs(
    y = "CPU RSS Memory (MB)",
    x = "Model Size",
    fill = NULL
  ) +
  theme_minimal()

gridExtra::grid.arrange(p3, p4, ncol = 2)

res_summary %>%
  kable(
    format  = "html",
    digits  = 1,
    caption = "GPU Peak & CPU RSS Memory by Model Size and Group"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE
  ) %>%
  column_spec(1, bold = TRUE)       %>%  # Bold the size column
  column_spec(2, width = "3cm")     %>%  # Wider group column
  column_spec(3:4, width = "2.5cm")      # Set width for numeric cols

```

### 2.8 Figure 7: Throughput & Latency Trends

```{r fig7-throughput-latency}
# Throughput & Latency by size & model group
tl_summary <- df_plot %>%
  group_by(size, group) %>%
  summarise(
    throughput_mean = mean(throughput),
    latency_mean    = mean(latency),
    .groups         = "drop"
  )

p_thr <- ggplot(tl_summary, aes(x = size, y = throughput_mean, color = group, group = group)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(
    title = "Mean Throughput by Model Group",
    x     = "Model Size",
    y     = "Throughput (examples/sec)"
  ) +
  theme_minimal()

p_lat <- ggplot(tl_summary, aes(x = size, y = latency_mean, color = group, group = group)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(
    title = "Mean Latency by Model Group",
    x     = "Model Size",
    y     = "Latency (sec/example)"
  ) +
  theme_minimal()

gridExtra::grid.arrange(p_thr, p_lat, ncol = 1)

tl_summary %>%
  kable(
    format  = "html",
    digits  = 2,
    caption = "Mean Throughput & Latency by Model Size and Group"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE
  ) %>%
  column_spec(1, bold = TRUE)   %>%  # Bold the size column
  column_spec(2, width = "3cm") %>%  # Wider group column
  column_spec(3:4, width = "2.5cm")   # Set width for numeric columns

```

### 2.9 Figure 8: Accuracy vs. Throughput Pareto

```{r fig8-pareto}
# Pareto scatter: Accuracy vs Throughput by model group
pareto_df <- df_plot %>%
  group_by(size, group) %>%
  summarise(
    acc_mean = mean(accuracy),
    thr_mean = mean(throughput),
    .groups  = "drop"
  )

ggplot(pareto_df, aes(x = thr_mean, y = acc_mean, color = group, shape = size)) +
  geom_point(size = 3) +
  labs(
    title = "Pareto: Accuracy vs Throughput",
    x     = "Mean Throughput (examples/sec)",
    y     = "Mean Accuracy",
    color = NULL
  ) +
  theme_minimal()

pareto_df %>%
  select(size, group, acc_mean, thr_mean) %>%
  kable(
    format  = "html",
    digits  = 3,
    caption = "Mean Accuracy vs Throughput by Model Size and Group"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE
  ) %>%
  column_spec(1, bold = TRUE)   %>%  # Bold the size column
  column_spec(2, width = "3cm") %>%  # Wider group column
  column_spec(3:4, width = "2.5cm")   # Width for numeric columns
```

### 2.10 Figure 9: Correlation Heatmap (4-bit FT)

```{r fig9-corr}
# Correlation heatmap for key metrics, by model group
# Use the full df_plot so it covers all runs
corr_mat <- df_plot %>%
  dplyr::select(accuracy, f1, throughput, latency, gpu_peak_mem_mb, cpu_rss_mb) %>%
  cor()

# Melt and plot
mm <- reshape2::melt(corr_mat)
ggplot(mm, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  viridis::scale_fill_viridis(name = "Correlation") +
  labs(
    title = "Correlation Heatmap of Key Metrics",
    x     = NULL,
    y     = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Convert to a data frame for printing
corr_df <- as.data.frame(corr_mat) %>%
  tibble::rownames_to_column(var = "Metric")

# Render as a styled HTML/PDF table
corr_df %>%
  kable(
    format  = "html",
    digits  = 2,
    caption = "Correlation Matrix of Key Metrics"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE
  ) %>%
  column_spec(1, bold = TRUE) %>%                # Bold the metric names
  column_spec(2:7, width = "2.5cm") 
```

### 2.11 Figure 10: Composite Cost-Effectiveness Frontier

```{r fig10-frontier}
# Composite Cost-Effectiveness Frontier by model group
frontier_df <- df_plot %>%
  group_by(size, group) %>%
  summarise(
    acc  = mean(accuracy),
    lat  = mean(latency),
    mem  = mean(gpu_peak_mem_mb + cpu_rss_mb),
    .groups = "drop"
  ) %>%
  mutate(
    # Normalize accuracy to [0,1]
    norm_acc  = (acc - min(acc)) / (max(acc) - min(acc)),
    # Define cost = latency + memory/10000, then normalize
    raw_cost  = lat + mem/10000,
    norm_cost = (raw_cost - min(raw_cost)) / (max(raw_cost) - min(raw_cost))
  )

ggplot(frontier_df, aes(x = norm_cost, y = norm_acc, color = group, label = size)) +
  geom_point(size = 4) +
  geom_text(nudge_y = 0.02) +
  labs(
    title = "Composite Cost-Effectiveness Frontier",
    x     = "Normalized Cost (lower is better)",
    y     = "Normalized Accuracy (higher is better)",
    color = NULL
  ) +
  theme_minimal()

frontier_df %>%
  select(size, group, acc, lat, mem, norm_acc, raw_cost, norm_cost) %>%
  kable(
    format  = "html",
    digits  = 3,
    caption = "Composite Cost-Effectiveness Frontier Data"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width        = FALSE
  ) %>%
  column_spec(1, bold = TRUE)   %>%  # Bold the size column
  column_spec(2, width = "3cm") %>%  # Wider group column
  column_spec(3:8, width = "2.5cm")   # Set width for numeric columns

```


